Automating Verification of MPC Security

### 2023-06-09 popl24/hyperprop.tex
checkpoint

The real/ideal model posits a \emph{simulator} that exists in an ideal
world, whose goal is to reconstruct observable information in a real
world run of a protocol $\prog$ that correctly implements an ideal
functionality $\idealf$- that is, execution of a configuration
$\config{\store}{\prog}$. We will assume that any protocol $\prog$
includes a public \emph{output view} $\outv$ where the result is
published, where $0$ is the ``public'' client and thus $\outv$
is of the form $\vx{0}{w}$ for some $w$, i.e., is owned by 0.
Cl
Note that the protocol is \emph{correct} iff the resulting
memory contains $\outv \mapsto \idealf(\store_S)$, where $S$ are the
secret inputs (honest and corrupt) in $\prog$.

The simulator is given just the corrupt inputs to the run- that is,
$\store_{S_C}$- and the output of the protocol- that is,
$\idealf(\store)$- and tries to reconstruct the \emph{distribution of
adversarial views} observed in the real world. The intuition here is
that the simulator is given the information that is necessarily
available to the adversary by observing a run of the protocl, and if
the simulator can reconstruct whatever information appears in the real
world adversarial views, by reconstructing their distributions, then
those views leak no additional information, i.e., they have no
additional dependencies on honest input secrets than what is revealed
by the output and conditionings on corrupt inputs.

@@

\subsection{Probabilistic Noninterference} Previous work has noted
relations between probabilistic noninterference and security
properties such as memory trace obliviousness \cite{XXX} and passive
security \cite{XXX}. And above (Section \ref{section-pmf-examples}) we
have demonstrated that our basic encryption scheme enjoys a noninterference
property. To formulate probabilistic noninterference in
our setting we begin by defining low equivalence of memories.
\begin{definition}[Memory Low Equivalence]
  Given $\store^1$ and $\store^2$ with $\dom(\store^1) = \dom(\store^2) = X$,
  we write $\store^1 =_C \store^2$ iff $\store^1_{X_C} = \store^2_{X_C}$.
\end{definition}
Now, we can define probabilistic noninterference as a property of
programs ensuring that low-equivalent initial memories produce
the same low-observable trace (i.e., corrupt views) \emph{with the
same probability}. 
\begin{definition}[Probabilistic Noninterference]
  Given a program $\prog$ with $\iov(\prog) = (S,V)$, we say that
  $\prog$ satisfies \emph{probabilistic noninterference} iff for all
  $\store_1, \store_2 \in \mems(S)$:
  $$\store_1 =_C \store_2 \implies
    \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}$$
\end{definition}

@@

\subsection{A Sound Hyperproperty for Passive Security}
The intuition that passive security means that the protocol releases
no more information than what is released by the output suggests a
refinement of noninterference. That is, the set of low equivalent
initial memories that can produce the same output should generate the
same distributions of adversarial views.
\begin{definition}[Noninterference Modulo Output]
  Given $H$, $C$, and a program $\prog$ with $\iov(\prog) = (S,V)$ and output
  view $x \in V$, we say that
  $\prog$ satisfies \emph{noninterference modulo output} iff for all
  $\store_1,\store_2 \in \mems(S)$:
  \begin{eqnarray*}
    & \store_1 =_C \store_2 \wedge
     (\margd{(\condd{\progd(\prog)}{\store_1})}{\{ x \}} =
     \margd{(\condd{\progd(\prog)}{\store_2})}{\{ x \}}) \\
    & \implies \\
    & \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}
  \end{eqnarray*}
\end{definition}
We can show, for example, that the 3-party xor protocol defined above
satisfies noninterference modulo output.

What is the formal connection with passive security? Note in
particular that the hyperproperty implies that the simulator can use
the protocol itself to reconstruct the distribution of adversarial
views by making a random guess about honest secrets given public
knowledge about the ideal functionality $\idealf$ a protocol
implements. However, there is a subtlety here in that the adversary
must be able to tractably reconstruct this knowledge. This is not
generally true, in particular if $\idealf$ is a 1-way
function. Define:
\begin{definition}[Ideal Knowledge]
  Given a functionality $\idealf$ and output value $v$, the associated
  \emph{ideal knowledge}, denoted $\ik(\idealf,v)$ is:
  $$
  \{ \store | \idealf(\store) = v \}
  $$
  We say that $\idealf$ is \emph{invertible} iff $\ik(\idealf, v)$ for all
  $v$ can be computed tractably.
\end{definition}
Now we can show our main result, that noninterference modulo output,
together with the assumption that $\idealf$ is invertible, implies
passive security of a protocol\footnote{We emphasize that the converse
is not asserted.}. 
\begin{theorem}
  Assume given invertible $\idealf$ and a protocol $\prog$ that
  correctly implements $\idealf$.  If program $\prog$ satisfies
  noninterference modulo output for all $H$ and $C$ with $|H| < |C|$,
  then $\prog$ is passive secure.
\end{theorem}

@@

  Now, since the random tape is selected in simulation from the same distribution
  that we assume for the real world, after selection of $\store'$ the
  probability that any particular $\store^{\SIM}_{V_C}$ is returned is by definition:
  $$
   (\margd{(\condd{\progd(\prog)}{\store'})}{V_C})(\store^{\SIM}_{V_C})
  $$
  Furthermore, since we assume that $\prog$ correctly implements $\idealf$, this
  means:
  $$
  \store_S =_C \store' \wedge
     (\margd{(\condd{\progd(\prog)}{\store_S})}{\{ x \}} =
      \margd{(\condd{\progd(\prog)}{\store'})}{\{ x \}})
  $$
  so by the definition of noninterference modulo output the choice
  any choice of $\store'$ yields the same distribution, i.e.:
  $$\margd{(\condd{\progd(\prog)}{\store_S})}{V_C}$$
  Thus, by definition:
  $$
   \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \margd{(\condd{\progd(\prog)}{\store_S})}{V_C}
  $$
  \qed
\end{proof}

>>>>

The real/ideal model posits a \emph{simulator} that exists in an ideal
world, whose goal is to reconstruct observable information in a real
world run of a protocol $\prog$ that correctly implements an ideal
functionality $\idealf$- that is, execution of a configuration
$\config{\store}{\prog}$. We will assume that any protocol $\prog$
includes a public \emph{output view} $\outv$ where the result is
published, where $0$ is the ``public'' client and thus $\outv$
is of the form $\view{0}{w}$ for some $w$, i.e., is owned by 0.
Note that the protocol is \emph{correct} iff the resulting
memory contains $\outv \mapsto \idealf(\store_S)$, where $S$ are the
secret inputs (honest and corrupt) in $\prog$.

The simulator is given just the corrupt inputs to the run- that is,
$\store_{S_C}$- and the output of the protocol- that is,
$\idealf(\store)$- and aims to reconstruct the distribution of
adversarial views observed in the real world. The intuition here is
that the simulator is given the information that is necessarily
available to the adversary by observing a run of the protocl, and if
the simulator can reconstruct whatever information appears in the real
world adversarial views, by reconstructing their distributions, then
those views leak no additional information, i.e., they have no
additional dependencies on honest input secrets than what is revealed
by the output and conditionings on corrupt inputs.

@@

\subsection{Probabilistic Noninterference}
\label{section-ni}
Previous work has noted
relations between probabilistic noninterference and security
properties such as memory trace obliviousness \cite{XXX} and passive
security \cite{XXX}. And above (Section \ref{section-pmf-examples}) we
have demonstrated that our basic encryption scheme enjoys a noninterference
property. To formulate probabilistic noninterference in
our setting we begin by defining low equivalence of memories.
\begin{definition}[Memory Low Equivalence]
  Given $\store^1$ and $\store^2$ with $\dom(\store^1) = \dom(\store^2) = X$,
  we write $\store^1 =_C \store^2$ iff $\store^1_{X_C} = \store^2_{X_C}$.
\end{definition}
Now, we can define probabilistic noninterference as a property of
programs ensuring that low-equivalent initial memories produce
the same low-observable trace (i.e., corrupt views) \emph{with the
same probability}. 
\begin{definition}[Probabilistic Noninterference]
  Given a program $\prog$ with $\iov(\prog) = (S,V)$, we say that
  $\prog$ satisfies \emph{probabilistic noninterference} iff for all
  $\store_1, \store_2 \in \mems(S)$:
  $$\store_1 =_C \store_2 \implies
    \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}$$
\end{definition}

@@

\subsection{A Sound Hyperproperty for Passive Security}
\label{section-nimo}
The intuition that passive security means that the protocol releases
no more information than what is released by conditioning on the
public output and corrupt inputs suggests a
refinement of noninterference. That is, the set of low equivalent
initial memories \emph{that can produce the same output} should generate the
same distributions of adversarial views.
\begin{definition}[Noninterference Modulo Output]
  We say that a program  $\prog$ satisfies \emph{noninterference modulo output},
  written $\NIMO(\prog)$, iff for all $H$, $C$, and 
  $\store_1,\store_2 \in \mems(S)$ we have:
  \begin{eqnarray*}
    & \store_1 =_C \store_2 \wedge
     (\margd{(\condd{\progd(\prog)}{\store_1})}{\{ \outv \}} =
     \margd{(\condd{\progd(\prog)}{\store_2})}{\{ \outv \}}) \\
    & \implies \\
    & \margd{(\condd{\progd(\prog)}{\store_1})}{V_C} =
    \margd{(\condd{\progd(\prog)}{\store_2})}{V_C}
  \end{eqnarray*}
where $\iov(\prog) = (S,V)$.
\end{definition}

From an information flow perspective, we can say that a $\NIMO$
protocol does not release information beyond what is declassified
via the public output. From a real/ideal perspective, we can
observe that $\NIMO$ implies that the simulator can use
the protocol itself to reconstruct the distribution of adversarial
views. That is, given the output of the ideal functionality
$\idealf(\store_S)$, the simulator can randomly choose any
$\store'$ such that $\store' =_C \store$ and $\idealf(\store'_S) =
\idealf(\store_S)$ and run the protocol with initial memory $\store'$-
$\NIMO$ guarantees ideal reconstruction of the real-world corrupt views. 
 However, there is a subtlety here in that the adversary
must be able to tractably reconstruct this knowledge. This is not
generally true, in particular if $\idealf$ is a 1-way
function. Define:
\begin{definition}[Ideal Knowledge]
  Given a functionality $\idealf$ and output value $v$, the associated
  \emph{ideal knowledge}, denoted $\ik(\idealf,v)$ is:
  $$
  \{ \store | \idealf(\store) = v \}
  $$
  We say that $\idealf$ is \emph{invertible} iff $\ik(\idealf, v)$ for all
  $v$ can be computed tractably.
\end{definition}
Now we can show our main result, that noninterference modulo output,
together with the assumption that $\idealf$ is invertible, implies
passive security of a protocol.
\begin{theorem}
  Assume given invertible $\idealf$ and a protocol $\prog$ that
  correctly implements $\idealf$.  If program $\prog$ satisfies
  noninterference modulo output for all $H$ and $C$ with $|H| < |C|$,
  then $\prog$ is passive secure.
\end{theorem}

@@

  Now, since the random tape is selected in simulation from the same distribution
  that we assume for the real world, after selection of $\store'$ the
  probability that any particular $\store^{\SIM}_{V_C}$ is returned is by definition:
  $$
   (\margd{(\condd{\progd(\prog)}{\store'})}{V_C})(\store^{\SIM}_{V_C})
  $$
  Furthermore, since we assume that $\prog$ correctly implements $\idealf$, this
  means:
  $$
  \store_S =_C \store' \wedge
     (\margd{(\condd{\progd(\prog)}{\store_S})}{\{ x \}} =
      \margd{(\condd{\progd(\prog)}{\store'})}{\{ x \}})
  $$
  so by the definition of noninterference modulo output the choice
  any choice of $\store'$ yields the same distribution, i.e.:
  $$\margd{(\condd{\progd(\prog)}{\store_S})}{V_C}$$
  Thus, by definition:
  $$
   \progd(\SIM_C(\store_{S_C},\idealf(\store))) = \margd{(\condd{\progd(\prog)}{\store_S})}{V_C}
  $$
\end{proof}
